---
word_document:
  fig_height: 7
  fig_width: 10
author: "Mark Greenwood"
date: ''
output: word_document
title: "Time Series HW 2 - Due Friday 9/9 at 6 pm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW 2

This will not be written in report style - just provide numbered answers to each question. Include your edited R code either in-line (using something like R-markdown - the newer option to compile directly to a Word document is quite nice for controlling formatting after compiling) or as an appendix to your homework.  

__You can not work with anyone that you worked with on the first homework.__ You can discuss it with old partners but should try work as much as possible with new collaborators.

You will get a 5% bonus on this homework if you do this homework in a group of 2 or 3 but not if it is any larger. If you discuss the assignment with others but turn in separate assignments, you need to document any discussions you had and how it impacted your answers - treat this like citing your sources.

We'll continue working with temperatures in degrees F in the column labeled MMXT (https://dl.dropboxusercontent.com/u/77307195/rawbozemandata.csv) for Bozeman's MSU weather station. 

Either use my code or your code to generate a `Year` variable that includes the fractional information on the month. For example, January would be 1900+0/12 and December would be 1900+11/12 in the `Year` variable.


1) Re-estimate and then use the output to write out the estimated model that incorporates the linear time trend and the month component from HW 1. Use indicator/dummy variable notation and define all model aspects.

2) What is the model for an observation in a January? July? Simplify the full model for each specific month to just provide a function of `Year`. 

3) Has the mean temperature changed differently in different months? Load the `car` package and make a plot using something like `scatterplot(MMXT~Year|Month,data=rawd1,legend.plot=T,smoother=F)`. Discuss the results in the plot.

4) Explore the same research question as in #3 by fitting a model with a `Year` by `Month` interaction (include main `effects` too). Generate and report an F-test for the interaction in the model using either `anova` or `Anova` from the `car` package. Write out a one-sentence conclusion that summarizes the results of the test including information on the distribution of the test statistic under the null hypothesis in that sentence.

5) Return to the additive model (`Year+Month`) and let's consider the potential evidence for specific differences in the pairs of months. We can test the 12 choose 2 comparisons, maintaining overall or "family-wise" error rates of, say, 5% by using Tukey's Honest Significant Difference. This is a little more complicated to use than in a simpler One-Way ANOVA model because we need to adjust for the linear trend. But the method is easily extended to handle more complicated models and those with multiple predictors using the same code you can use in the One-Way ANOVA case (See Greenwood and Banner (2016) Section 2.5 and 2.6 (https://scholarworks.montana.edu/xmlui/handle/1/2999) for more on this method in the simpler scenario). Your code will be something like the following, after you install the `multcomp` package. Note that you may need to modify the model name (`model1` below) and the variable to perform Tukey's HSD on (`Monthf` below).

`require(multcomp)`

`Tukey_results<-glht(model1,linfct=mcp(Monthf="Tukey"))`

`summary(Tukey_results)`

`confint(Tukey_results)`

  You can also plot the results but with so many comparisons, the plot is not very useful. Based on these results, what can you say about evidence for differences in the months after controlling for the linear year trend? A fairly simple pattern should arise in the results. I find it useful to switch to saying differences were or were not "detected" when dealing with a large suite of Tukey's results. 

6) One other issue that wasn't obvious in either the initial plots or in the model diagnostics is whether the linear trend really is a good description of the changes over time. There are a couple of ways to add polynomial terms to linear models. The simplest is just to create a squared version of the variable and include it as an additional variable. We'll see that this isn't an optimal choice in the next homework, but let's start this way for now. Once you are considering incorporating polynomials, we need to consider some sort of model refinement to decide on the polynomial order (linear, quadratic, cubic, quartic, etc.) to use. One technique is to sequentially add higher order versions of the variable to the lower order (linear, quadratic, etc.) versions of it and stop when the highest order term has a "large p-value" and drop back to the next lower order model that has the highest order term having a small p-value (checking diagnostics to make sure nothing really high order was missed). Employ this approach, starting with adding a quadratic `Year` variable to the `Year+Month` model. Report the test for the quadratic component (with distribution, test statistic, and p-value) and interpret this result, remembering that t-tests are conditional on other stuff in the model. 

7) For the model with the linear and quadratic `Year` components and an additive `Month`, produce the 2x2 diagnostic plots from `plot(model)` as in HW 1. This time, discuss the top left panel (`Residuals vs Fitted`). Discuss what you can generally assess in this plot and then discuss the specific results for this model. 



8) Run the following code so I can see what version of R you are using:

### Documenting R version 

```{r}
getRversion()
```




